Who Needs Privacy?

### 2019-02-26 manuscript/manuscript.RmD
Major updates on everything. Specifically, this includes updates on data, including MMs latest feedback, and getting things ready for submission.

abstract          : "This study analyzes how personality traits relate to need for privacy. We focus on three dimensions: (a) need for privacy from government surveillance, (b) need for privacy from other individuals, and (c) need for anonymity. Using an online questionnaire with 271 student respondents, several significant relationships were found. For example, results showed that less sociable people desired considerably more privacy on all three dimensions. Somewhat more controversially, people who reported less integrity also desired noticeably more anonymity. While anxious people desired slightly less privacy from government surveillance, risk avoidant people desired considerably more privacy from other people. Traditionalism did not relate to need for privacy. Together, the results have several implications. For example, they shed some light on the nothing-to-hide argument: A person who desires more anonymity might indeed be of less integrity. However, it is equally plausible that this person is just a little less sociable."
keywords          : "Privacy, need for privacy, personality, anonymity, integrity, SEM"

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

@@

```{r setup, include = FALSE}
# devtools::install_git("https://github.com/tdienlin/td")
packages <- c("devtools", "english", "ggplot2", "kableExtra", "lavaan", "magrittr", 
              "papaja", "semPlot", "td", "tidyverse")
lapply(packages, library, character.only = TRUE)
load("../data/workspace.RData")

@@

In light of the increasing digitization of everyday life, which has led to several sweeping societal changes such as the commodification and monetization of personal information [@Sevignani.2016], privacy has become a major topic of interest. Despite its importance, however, we still know surprisingly little about the relation between privacy and personality [e.g., @Masur.2018, p. 155]. Why do some people need more privacy than others, and how do these people differ from one another? We think that a better understanding of the relation between personality and privacy is crucial: Several theories argue that personality determines privacy behaviors [e.g., @Masur.2018], however, to date there is almost no empirical research that can be used to deduct well-informed hypotheses. As a result, the main question of this paper is as follows: What are personality facets that can be used to best explain peoples' need for privacy?

@@

Privacy as a theoretical concept is both complicated and contested [@Nissenbaum.2010, p. 71]. Hence, we begin by outlining our own understanding of privacy. First and foremost, privacy captures the extent of voluntary withdrawal from others [@Westin.1967]. Several models suggest that privacy is multi-dimensional. For example, in a theory-driven treatise @Burgoon.1982 argued that privacy has four dimensions: informational, social, psychological, and physical privacy. @Pedersen.1979, by contrast, did an empirical factor analysis (initially starting with 94 items) and suggested that privacy exists on six dimensions: reserve, isolation, solitude, intimacy with friends, intimacy with family, and anonymity. In addition, @Schwartz.1968 differentiated between horizontal and vertical privacy; whereas horizontal privacy captures withdrawal from peers, vertical privacy addresses withdrawal from superiors or institutions (e.g., government agencies). Finally, one can also distinguish between the objective privacy context, the subjective privacy perception, the psychological need for privacy (which is both a situational and dispositional need), and the resulting privacy behavior (often represented by self-disclosure) [@Dienlin.2014].

For the purpose of this study, we combine the aforementioned notions and focus on (a) vertical privacy with regard to the need for withdrawal from government surveillance, (b) horizontal privacy in terms of the need for withdrawal from peers, friends, or acquaintances, and (c) both horizontal and vertical privacy as captured by the general need for anonymity.

In order to guide the selection of personality dimensions that might explain need for privacy best, it is useful to determine why people actually need privacy. @Westin.1967 defined four primary positive purposes of privacy: (1) self-development (i.e., the integration of experiences into meaningful patterns), (2) autonomy (i.e., the desire to avoid being manipulated and dominated), (3) emotional release (i.e., the release of tension from social role demands), and (4) protected communication (i.e., the ability to foster intimate relationships). Not least, privacy facilitates self-disclosure [@Dienlin.2014], which is vital for attaining social support, initiating relationships, and getting close to one another [@Omarzu.2000].

On the other hand, however, too much privacy can also be problematic. Because human beings are inherently social, being overly cut-off from others can impede flourishing, nurture deviant behavior, or introduce power asymmetries [@Altman.1975]. "[T]he feminist critique rightfully argued that for men, the private sphere can be a shielded sphere in which female suppression is not sanctioned." [@DeCew.1997, p. 81, as cited by Masur, 2018, p. 38]. Next, and as mentioned above, privacy fosters self-disclosure. However, any self-disclosure is also always a potential risk because others might disagree, disapprove, or misuse the information in other contexts [@Petronio.2010].

@@

In what follows we now explore several personality aspects that we think might predict the need for privacy. As there is no established theory that dovetails privacy and personality, it is difficult to deduct precise and well-informed hypotheses. The rationales that we outline below should hence considered to be tentative suggestions. Also, please note that we elaborate only on those dimensions of need for privacy for which plausible rationales could be found.

The general underlying theoretical framework that guided our selection process was the big five approach [e.g.,@John.1999]. Note that privacy concerns, a variable conceptually close to need for privacy, shows only negligible relations with the big five factors [@Bansal.2010; @Junglas.2008]. In order to be more precise, we hence follow the advice by @Paunonen.2001 and refer to specific personality facets instead of generic personality factors. To illustrate, instead of dwelling on the general factor of _extraversion_, we instead focused on the facet _gregariousness_, which we consider to be more pertinent.

Our reasoning was guided by another central theoretical thought. As implied above, privacy can be either positive or negative. Similarly, other people, the government, and anonymity can be considered either a _resource_ or a _threat_. Having information about a person's personality can inform us whether he or she is more likely to think of others as a resource or a threat. It follows that if other people are considered a threat it seems to be more likely that a person will desire more privacy from others, and vice versa [@Altman.1976].

@@

First, we argue that need for privacy should be closely related to a person's sociability or gregariousness [which is a subdimension of extraversion, @John.1999]. Sociability captures whether people prefer to spend their time alone or with company. It seems plausible that people who are more sociable are also more likely to think of other people as a resource, which is why they should generally desire less interpersonal privacy and less anonymity [e.g., @Buss.2001]. Put differently, given that privacy is a voluntary withdrawal from society [@Westin.1967], we expect that people who are less sociable, more reserved, or more shy should have a greater need for privacy from others.

This rationale is supported by several empirical studies: Extroverted people desire less privacy [@Morton.2013], people who describe themselves as introverted thinkers are more likely to prefer social isolation [@Pedersen.1982], and introverted people are more likely to report invasions of privacy [@Stone.1986].

@@

More controversially, it has been argued that people need privacy because they have something to hide. The so-called nothing-to-hide argument states that “If you have nothing to hide, you have nothing to fear.” As described by Solove, the nothing-to-hide argument says that data mining and surveillance by government entities “is not likely to be threatening to the privacy of law-abiding citizens. Only those who are engaged in illegal activities have a reason to hide this information” [@Solove.2007, p. 753]. Hence, another potential predictor of why people need privacy could also be a so-called "lack of integrity".

Because integrity is a delicate concept, let us first try to define it conceptually. Although in terms of a scientific definition there is no consensus, most scholars seem to agree that integrity “incorporates a tendency to comply with social norms, avoid deviant behavior, and embrace a sense of justice, truthfulness, and fairness” [@Connelly.2006, p. 82]. In order to sidestep the (very legitimate) philosophical debates about what constitutes integrity and what not, we hence follow @Paunonen.2002 and adopt a lowest common denominator approach, which means that we only consider participating in explicitly socially-sanctioned or illegal activities as a sign of lack of integrity.

It is indeed possible to think of theoretical arguments for why lack of integrity might correlate positively with need for privacy. People who have actually committed crimes evidently face even greater risk from self-disclosure compared to others, because government agencies and people sould surely disapprove of their activities [@Petronio.2010]. Hence, the government and other people are more likely to be a threat, which should render anonymity rather a resource. As a consequence, people with lower integrity might desire more privacy as a means to mitigate their felt risk [@Altman.1976].

There are also a few empirical studies that imply---at least indirectly---a relation between privacy and integrity. For example, studies have found that surveillance can reduce cheating behaviors [@Corcoran.1987; @Covey.1989]. @Covey.1989 for example asked students to solve an impossible maze. In the high surveillance condition, the experimenter stood in front of the students and closely monitored their behavior. In the low surveillance condition, the experimenter remained behind the students where he or she could not see the students. Results showed greater cheating among students in the low surveillance condition, suggesting that in situations with less privacy, people show more integrity (i.e., fewer cheating behaviors). Next, in a longitudinal sample with 457 respondents in Germany [@Trepte.2013a], people who felt they needed more privacy were also less authentic on their SNSs profiles (_r_ = -.48) and less authentic in their personal relationships (_r_ = -.28). Given the argument that authenticity is a subset of integrity [@Sheldon.2004], one could hence reason that also the concept of integrity might relate to a person’s need for privacy. Somewhat related, it has been found that people who are more agreeable are also moderately less concerned about their privacy [@Junglas.2008]. Finally, @Pedersen.1982 showed that three dimensions of need for privacy relate to self-esteem: Respondents who held a lower self-esteem were more reserved (_r_ = .29), needed more anonymity (_r_ = .21), and preferred solitude (_r_ = .24). While self-esteem and integrity are distinct concepts, Pedersen’s specific operationalization of self-esteem integrated several aspects of integrity by using items such as “moral, nice, fair, unselfish, good, honest, reputable, sane”.

In conclusion, we could imagine that lack of integrity may indeed relate to an increased need for privacy from government surveillance, as governments have the legitimate power to prosecute illegal activities. Next, it seems plausible that lack of integrity relates to an increased need for anonymity, as anonymity makes it more difficult for both government and social agents to identify and address potential wrongdoers. Finally, lack of integrity may also relate to an increased need for privacy from other people, as most other people will disapprove of immoral or illegal activities, and could even reveal those activities to authorities.

@@

Next, it seems possible that the need for privacy is also related to people's level of general anxiety [which is a subdimension of neuroticism, @John.1999]. General anxiety measures for example whether people are afraid of negative events or whether they are easily frightened. With regard to interpersonal privacy, one could argue that people who are anxious are more likely to feel that social interactions pose a greater risk and threat [especially with strangers or weak ties; @Granovetter.1973], which is why anxious people might desire more privacy. Somewhat related, prior empirical research has shown that people who are more concerned about their privacy are also more likely to withdraw online, for example by deleting posts or untagging themselves from linked content [@Dienlin.2016a]. On the other hand, one could argue in favor of the opposite: People who are more anxious may desire _less_ privacy from others (especially their strong ties), as a means to cope better with their daily challenges.

Concerning the need for privacy from government surveillance, we could imagine that people who are more anxious desire less privacy. Despite the fact that only 18% of all US-Americans trust their government "to do what is right" [@PewResearchCenter.2017], almost everyone agrees that "it's the government's job to keep the country safe", with most people also being satisfied with the government's job [@PewResearchCenter.2015c]. Hence, when being anxious, the government should rather be seen as a resource than a threat. It therefore seems plausible that people who are in general more anxious are also more likely to consent to government surveillance, given that these programs promise to prevent crime or to reduce the likelihood of terrorist attacks [@Greenwald.2013], implying that people who are more anxious might desire less privacy from government surveillance and, for the same reasons, also less anonymity.

### Risk aversion

Similarly, it could be that need for privacy is related to the general tendency to avoid risks, to deliberate, and to plan carefully [deliberation is a subdimension of conscientiousness, @John.1999]. Risk avoidance captures the degree to which people prefer to abstain from taking risks. If other people are considered a threat, people who are risk averse should desire more interpersonal privacy, because they may feel greater risk associated with disclosure of personal information. The most cautious strategy to minimize risks of information disclosure would be, arguably, to keep as much information as possible private. Somewhat related, empirical studies report that people who think that their privacy is at risk are less likely to disclose information online [e.g., @Bol.2018]. Moreover, research suggests that conscientious people are slightly more concerned about their privacy [@Junglas.2008]. But again, especially with regard to privacy from government surveillance, risk averse people could also desire _less_ privacy, in order for the government to be able to avert potential threats.

@@

Next, it seems plausible that need for privacy is also related to traditionality [which is a subdimension of openness to experience, @John.1999]. Traditionalism measures whether people prefer to stick with their usual routines. The computers and the Internet have rendered the world increasingly knowable: Social interactions, purchases, and medical treatments nowadays all produce digital traces, which can be combined into accurate latent user profiles. Given that digital information is persistent, searchable, reproducible, and scalable [@boyd.2008c], this allows for unprecedented means and degrees of surveillance. Mark Zuckerberg famously observed that privacy is no longer a social norm, rather the norm is that people share personal information [@Johnson.2010]. Hence, in order to be part of contemporary life, it seems necessary to give up some privacy. However, people may not be equally willing to pay that price, and especially people who are more traditional might decide against giving up their privacy.

Empirical research does find that older people, who are generally less open and more traditional [@Donnellan.2008], are more concerned about their privacy [@Fife.2012]. On the other hand, @Junglas.2008 report that openness to experience is positively related to privacy concern, which would argue in favor of the exact opposite. Taken together, we still consider it plausible that people who are more traditional also desire more privacy in all three aforementioned contexts.

@@

Finally, it has often been shown that socio-demographic variables such as sex, age, and affluence can relate to the need for privacy [@Park.2015; @Tifferet.2019; @Weinberger.2017b; @Trepte.2013a]. Although these variables do not specifically address our research question at hand they will be included---not least as their omission could lead to spurious results or an inflation of false positives.

@@

First, we tested the factorial validity of the variables we collected. In a first step, we ran confirmatory factor analyses (CFAs) to determine whether the variables were uni-dimensional. If fit was not adequate, we then ran exploratory factor analyses (EFAs) to assess the actual underlying factor structure. If more than one factor was revealed, we tried to find a bifactor model solution.[^1] This way we were able to attain a general measure of the predictors, instead of needing to further complicate our analyses by including subdimensions of the predictors.[^2] If no adequate bifactor model could be found, we proceeded by deleting items with low loadings on the general factor and/or the specific factors. If also after deletion of individual items no bifactor solution emerged, using a subset of the items we then aimed to extract a single factor with sufficient factorial validity.

@@

To anwers our research question, we first had a look at the variables' bivariate relations (see Figure \@ref(fig:bivar-fig)). To better assess the variables' potential causal relations, we next ran a multivariate structural regression model. Because we were interested in a complex model (overall, eight predictors and three outcomes) but the sample was comparatively small, we simplified the model. To this end, instead of a fully latent structural regression model we conducted a partially latent structural regression model, in which the predictor variables were modelled as single indicators while controlling for measurement error [@Kline.2016, p. 214]. To get high-quality single indicators of the predictors, we computed the average of the model predicted values / latent factor scores, which we extracted from the CFAs. If the CFAs showed a single-factor solution, we used the model predicted values for this latent factor; if the CFAs produced a bifactor solution, we used the model predicted values for the general latent factor. As main variable of interest, the outcome need for privacy was measured as a three-dimensional latent factor consisting of overall 12 items (see below).

Finally, combining the information of several items into a latent factors helps to reduce and combine information. At the same time, however, this approach also introduces several researcher degrees of freedom. We hence emphasize that we deleted items only in order to achieve satisfactory factorial validity and not in order to cherry-pick significant material. In light of our not having preregistered the analyses and in order to provide the complete picture, in the online supplementary material (OSM) we hence also report how the independent variables predicted each item individually.

`r cite_r("../bibliography/r-references.bib", withhold = FALSE, pkgs = c("GGally", "ggplot2", "knitr", "lavaan", "mice", "MVN", "papaja", "PerFit", "psych", "pwr", "semTools", "tidyverse"))` was used for the analyses.

@@

As the main variable of interest, we collected several items to measure need for privacy. First, we collected four items capturing the need for informational privacy using the scale developed by @Trepte.2017d. Second, we measured need for privacy on a societal level using nine self-designed items. The first subdimension was government surveillance, which represents the extent to which people want the government to abstain from collecting information about them. The second dimension was anonymity, which captures the extent to which people feel the need to avoid identification. Third, we measured need for privacy on an interpersonal level using nine self-designed items. The first subdimension measures need for privacy from other people in online contexts. The second subdimension measures need for privacy from other people in offline contexts.

EFAs and CFAs revealed a well-fitting model with 12 items (one double-loading), which consisted of the three dimensions (a) need for privacy from the government, (b) need for privacy from other people, and (c) need for anonymity. For a list of all items, see Table \@ref(tab:need-for-privacy-items).

```{r need-for-privacy-items}
d_t_items <- rbind(
  t_pri_nee %>% filter(name %in% pri_nee_gov_items),
  t_pri_nee %>% filter(name %in% pri_nee_int_items),
  t_pri_nee %>% filter(name %in% pri_nee_ano_items)) %>% 
  select("Item content" = content) %>% 
  mutate_all(linebreak)
kable(d_t_items, "latex", booktabs = TRUE, linesep = "",
      caption = "Overview of Items Measuring Need for Privacy") %>% 
  # column_spec(1, width = "16cm") %>% 
  kable_styling(latex_options="scale_down") %>% 
  group_rows("Privacy need government", 1, 5) %>%
  group_rows("Privacy need interpersonal", 6, 9) %>%
  group_rows("Privacy need anonymity", 10, 13)
```

@@

Sociability was measured with the extraversion subscale _gregariousness_ [@John.1999], which consists of eight items. An example item is “I shy away from crowds of people.” Analyses revealed that a bifactor model based on a subset of six items showed good fit to the data.

@@

Anxiety was measured with the neuroticism subscale _anxiety_ [@John.1999], which consists of eight items. An example item is “I am easily frightened.” Analyses revealed that a bifactor model using all eight items showed good fit to the data.

>>>>

abstract          : 'This study analyzes how personality traits relate to need for privacy. We focus on three dimensions: (a) need for privacy from government surveillance, (b) need for privacy from other individuals, and (c) need for anonymity. Using an online questionnaire with 273 student respondents, several significant relationships were found. For example, results showed that less sociable people desired considerably more privacy on all three dimensions. Somewhat more controversially, people who reported less integrity also desired more anonymity. While anxious people desired slightly less privacy from government surveillance, risk avoidant people desired considerably more privacy from other people. Traditionalism did not relate to need for privacy. Together, the results have several implications. For example, they shed light on the "nothing-to-hide argument": A person who desires more anonymity might indeed be of less integrity. The most important predictor of need for privacy, however, was sociability. In conclusion, this study is the first to show that if a person desires more privacy, it is most plausible that he or she is less sociable.'
keywords          : "Privacy, need for privacy, personality, anonymity, integrity, SEM"

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
\ output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

@@

```{r setup, include = FALSE}
devtools::install_github("tdienlin/td@v.0.0.2.5")
packages <- c("devtools", "english", "ggplot2", "kableExtra", "lavaan", 
              "magrittr", "papaja", "semPlot", "td", "tidyverse")
lapply(packages, library, character.only = TRUE)
load("../data/workspace.RData")

@@

In light of the increasing digitization of everyday life, which has led to several sweeping societal changes such as the commodification and monetization of personal information [@Sevignani.2016], privacy has become a major topic of societal interest. Despite its importance, however, we still know surprisingly little about the relation between privacy and personality [e.g., @Masur.2018, p. 155]. Why do some people feel they need or desire more privacy than others do, and how do these people differ from one another? We think that a better understanding of the relation between personality and privacy is crucial: Several theories argue that personality determines privacy behaviors [e.g., @Masur.2018], however, to date there is almost no empirical research that can be used to deduce well-informed hypotheses. As a result, the main question of this paper is: What are personality facets that can be used to best explain peoples' felt need for privacy?

@@

Privacy as a theoretical concept is both complicated and contested [@Nissenbaum.2010, p. 71]. Hence, we begin by outlining our own understanding of privacy. First and foremost, privacy captures the extent of voluntary withdrawal from others [@Westin.1967]. Several models suggest that privacy is multi-dimensional. For example, in a theory-driven treatise @Burgoon.1982 argued that privacy has four dimensions: informational, social, psychological, and physical privacy. @Pedersen.1979, by contrast, argues that on the basis of an empirical factor analysis (initially starting with 94 items) privacy exists on six dimensions: reserve, isolation, solitude, intimacy with friends, intimacy with family, and anonymity. In addition, @Schwartz.1968 differentiated between horizontal and vertical privacy; whereas horizontal privacy captures withdrawal from peers, vertical privacy addresses withdrawal from superiors or institutions (e.g., government agencies). The need for privacy, then, captures "[a]n individual’s need to selectively control the
access of others to the individual = self with the aim of achieving a desired level of physical or psychological privacy [...]") [@Trepte.2017d, p. 1]. Finally, one can also distinguish between the objective privacy context (e.g., how many users can actually read a post on an SNS?), subjective privacy perceptions (e.g., how many users people _think_ can read a post on an SNS?), and resulting privacy behaviors (e.g., how much do people self-disclose?) [@Dienlin.2014].

For the purpose of this study, we combine the aforementioned notions and focus on (a) vertical privacy with regard to people's felt need for withdrawal from government surveillance, (b) horizontal privacy in terms of the perceived need for withdrawal from peers, friends, or acquaintances, and (c) both horizontal and vertical privacy as captured by people's general felt need for anonymity.

In order to guide the selection of personality dimensions that might explain need for privacy best, it is useful to determine why people actually feel they need privacy. According to @Trepte.2017d, the need for privacy is a secondary need---it is not an end in itself, but rather a way to satisfy other more fundamental needs. Specifically, @Westin.1967 defined four ultimate purposes of privacy: (1) self-development (i.e., the integration of experiences into meaningful patterns), (2) autonomy (i.e., the desire to avoid being manipulated and dominated), (3) emotional release (i.e., the release of tension from social role demands), and (4) protected communication (i.e., the ability to foster intimate relationships). Not least, privacy facilitates self-disclosure [@Dienlin.2014], which is vital for attaining social support, initiating relationships, and getting close to other people [@Omarzu.2000].

On the other hand, however, there are negative aspects of privacy too. For example, too much privacy can be problematic. Because human beings are inherently social, being overly cut-off from others can impede flourishing, nurture deviant behavior, or introduce power asymmetries [@Altman.1975]. Also, and as mentioned above, privacy fosters self-disclosure, which presents a potential risk because others might disagree, disapprove, or misuse the information in other contexts [@Petronio.2010]. The dialectical tension between the positive and negative aspects of privacy likely causes variability across individuals in their need for privacy. This study asks: What role does personality play in determining individual-level variations in need for privacy?

@@

In what follows we now explore several personality aspects that we think might predict the need for privacy. As there is no established theory that dovetails privacy and personality, it is difficult to deduce precise and well-informed hypotheses. The rationales that we outline below should hence be considered tentative suggestions toward a theory of personality and privacy. Also, please note that we elaborate only on those dimensions of need for privacy for which plausible rationales could be formed.

The general underlying theoretical framework that guided our selection process was the Big Five approach [e.g., @John.1999]. Note that privacy concerns, a variable that is conceptually close to need for privacy, shows only negligible relations with the Big Five factors [@Bansal.2010; @Junglas.2008]. In order to be more precise, we hence follow the advice by @Paunonen.2001 and refer to specific personality facets instead of generic personality factors. To illustrate, instead of dwelling on the general factor of _Extraversion_, we instead focused on the facet _Gregariousness_, which we consider to be more pertinent.

Our reasoning further was guided by another central theoretical tenet. As suggested above, privacy can be either positive or negative. Similarly, other people, the government, and anonymity can be considered either a _resource_ or a _threat_. Having information about a person's personality can inform us whether he or she is more likely to think of others as a resource or a threat. It follows that if other people are considered a threat it seems to be more likely that a person will desire more privacy from others, and vice versa [@Altman.1976].

@@

First, we argue that need for privacy should be closely related to a person's sociability or gregariousness [which is a subdimension of extraversion, @Costa.1992b]. Sociability captures whether people prefer to spend their time alone or with company. It seems plausible that people who are more sociable are also more likely to think of other people as a resource, which is why they should generally desire less interpersonal privacy and less anonymity [e.g., @Buss.2001]. Put differently, given that privacy is a voluntary withdrawal from society [@Westin.1967], we expect that people who are less sociable, more reserved, or more shy should have a greater need for privacy from others.

This rationale is supported by several empirical studies: People who score higher on the personality meta-factor _plasticity_, which is a composite of the two personality factors extraversion and openness, have been shown to desire less privacy [@Morton.2013]; people who describe themselves as introverted thinkers are more likely to prefer social isolation [@Pedersen.1982]; and introverted people are more likely to report invasions of privacy [@Stone.1986].

@@

More controversially, it has been argued by some that people need privacy because they have something to hide. The so-called nothing-to-hide argument states that “If you have nothing to hide, you have nothing to fear.” As described by Solove, the nothing-to-hide argument says that data mining and surveillance by government entities “is not likely to be threatening to the privacy of law-abiding citizens. Only those who are engaged in illegal activities have a reason to hide this information” [@Solove.2007, p. 753]. Hence, another potential predictor of why people need privacy could also be a so-called "lack" of integrity.

Because integrity is a delicate concept, let us first try to define it conceptually. Although in terms of a scientific definition there is no consensus, most scholars seem to agree that integrity “incorporates a tendency to comply with social norms, avoid deviant behavior, and embrace a sense of justice, truthfulness, and fairness” [@Connelly.2006, p. 82]. In order to sidestep the (very legitimate) philosophical debates about what constitutes integrity and what not, we hence follow @Paunonen.2002 and adopt a lowest common denominator definition, which means that we only consider participating in explicitly socially-sanctioned or illegal activities as a sign of lack of integrity.

It is possible to think of theoretical arguments for why lack of integrity might correlate positively with a person's felt need for privacy. People who actually commit crimes may face even greater risk from self-disclosure compared to others, because government agencies and people would surely disapprove of their activities [@Petronio.2010]. Hence, the government and other people are more likely to be perceived as a threat, which should render anonymity  a resource. As a consequence, people with lower integrity might desire more privacy as a means to mitigate their felt risk [@Altman.1976].

There are also a few empirical studies that imply---at least indirectly---a relation between privacy and integrity. For example, studies have found that surveillance can reduce cheating behaviors [@Corcoran.1987; @Covey.1989]. @Covey.1989 for example asked students to solve an impossible maze. In the high surveillance condition, the experimenter stood in front of the students and closely monitored their behavior. In the low surveillance condition, the experimenter remained behind the students where he or she could not see the students. Results showed greater cheating among students in the low surveillance condition, suggesting that in situations with less privacy, people show more integrity (i.e., fewer cheating behaviors). Next, in a longitudinal sample with 457 respondents in Germany [@Trepte.2013a], people who felt they needed more privacy were also less authentic on their online social network profiles (_r_ = -.48) and less authentic in their personal relationships (_r_ = -.28). Given the argument that authenticity is a subset of integrity [@Sheldon.2004], one could hence also reason that the concept of integrity might relate to a person’s perceived need for privacy. Somewhat related, it has been found that people who are more agreeable are also moderately less concerned about their privacy [@Junglas.2008]. Finally, @Pedersen.1982 showed that three dimensions of need for privacy relate to self-esteem: Respondents who held a lower self-esteem were more reserved (_r_ = .29), needed more anonymity (_r_ = .21), and preferred solitude (_r_ = .24). While self-esteem and integrity are distinct concepts, Pedersen’s specific operationalization of self-esteem integrated several aspects of integrity by using items such as moral, nice, fair, unselfish, good, honest, and reputable (p. 12).

In conclusion, we could imagine that lack of integrity may indeed relate to an increased felt need for privacy from government surveillance, as governments have the legitimate power to prosecute illegal activities. Next, it seems plausible that lack of integrity may relate to an increased need for anonymity, as anonymity makes it more difficult for both government and social agents to identify and address potential wrongdoers. Finally, lack of integrity may also relate to an increased need for privacy from other people, as most other people will disapprove of immoral or illegal activities, and could even reveal those activities to authorities.

@@

Next, it seems possible that the need for privacy is also related to people's level of general anxiety [which is a subdimension of neuroticism, @Costa.1992b]. General anxiety measures for example whether people are afraid of negative events or whether they are easily frightened. With regard to interpersonal privacy, one could argue that people who are anxious are more likely to feel that social interactions pose a greater risk and threat [especially with strangers or weak ties; @Granovetter.1973], which is why anxious people might desire more privacy. Somewhat related, prior empirical research has shown that people who are more concerned about their privacy are also more likely to withdraw online, for example by deleting posts or untagging themselves from linked content [@Dienlin.2016a]. On the other hand, one could argue in favor of the opposite: People who are more anxious may desire _less_ privacy from others (especially their strong ties), as a means to cope better with their daily challenges.

Concerning the need for privacy from government surveillance, we could imagine that people who are more anxious desire less privacy. Despite the fact that only 18% of all Americans trust their government "to do what is right" [@PewResearchCenter.2017], almost everyone agrees that "it's the government's job to keep the country safe," with most people also being satisfied with the government's job [@PewResearchCenter.2015c]. Hence, for anxious individuals, the government might be seen as a resource rather than a threat. It therefore seems plausible that people who are in general more anxious are also more likely to consent to government surveillance, given that such surveillance promises to prevent crime or to reduce the likelihood of terrorist attacks [@Greenwald.2013], implying that people who are more anxious might desire less privacy from government surveillance and, for the same reasons, also less anonymity.

Similarly, it could be that an individual's felt need for privacy is related to a general tendency to avoid risks, to deliberate, and to plan carefully [deliberation is a subdimension of conscientiousness, @Costa.1992b]. Risk avoidance captures the degree to which people prefer to abstain from taking risks. If other people are considered a threat, people who are risk averse should desire more interpersonal privacy, because they may feel greater risk associated with disclosure of personal information. The most cautious strategy to minimize risks of information disclosure would be, arguably, to keep as much information as possible private. Somewhat related, empirical studies report that people who think that their privacy is at risk are less likely to disclose information online [e.g., @Bol.2018]. Moreover, research suggests that conscientious people are slightly more concerned about their privacy [@Junglas.2008]. But as above, especially with regard to privacy from government surveillance, risk averse people could also desire _less_ privacy, in order for the government to be able to avert potential threats. In sum, think that it is most plausible that people who are more risk averse also desire more privacy in all three contexts measured in this study.

@@

Next, it seems plausible that need for privacy is also related to traditionality [which is a subdimension of openness to experience, @Costa.1992b]. Traditionalism measures whether people prefer to stick with their usual routines. Computers and the Internet have rendered the world increasingly knowable: Social interactions, purchases, and medical treatments nowadays all produce digital traces, which can be combined into accurate latent user profiles. Given that digital information is persistent, searchable, reproducible, and scalable [@boyd.2008c], this allows for unprecedented means and degrees of surveillance. Mark Zuckerberg famously observed that privacy is no longer a social norm, rather the norm is that people share personal information [@Johnson.2010]. Hence, in order to be part of contemporary life, it seems necessary to give up some privacy. However, people may not be equally willing to pay that price, and especially people who are more traditional might decide against giving up their privacy.

Empirical research does find that older people, who are generally less open and more traditional [@Donnellan.2008], are more concerned about their privacy [@Fife.2012]. On the other hand, @Junglas.2008 report that openness to experience is positively related to privacy concern, which would argue in favor of the exact opposite pattern of results. Taken together, we still consider it plausible that people who are more traditional also desire more privacy in all three contexts measured in this study.

@@

Finally, it has often been shown that socio-demographic variables such as sex, age, and affluence can relate to the need for privacy [@Park.2015; @Tifferet.2019; @Weinberger.2017b; @Trepte.2013a]. Although these variables do not specifically address our research question at hand they will be included as control variables---not least as their omission could lead to spurious results or an inflation of false positives.

@@

First, we tested the factorial validity of the variables we collected. In a first step, we ran confirmatory factor analyses (CFAs) to determine whether the variables were unidimensional. If fit was not adequate, we then ran exploratory factor analyses (EFAs) to assess the actual underlying factor structure. EFAs were run using maximum likelihood estimation and oblimin rotation [@Osborne.2004b, p. 7]. If more than one dimension was revealed, we tried to find a bifactor model solution.[^1] This way we were able to attain a general measure of the predictors, instead of needing to further complicate our analyses by including subdimensions of the predictors.[^2] If no adequate bifactor model could be found, we proceeded by deleting items with low loadings on the general factor and/or the specific factors. If also after deletion of individual items no bifactor solution emerged, we then aimed to extract a single factor with sufficient factorial validity using a subset of the items.

@@

To anwers our research question, we first had a look at the variables' bivariate relations (see Figure \@ref(fig:bivar-fig)). To better assess the variables' potential causal relations, we next ran a multivariate structural regression model. Because we were interested in a complex model (overall, eight predictors and three outcomes) but the sample was comparatively small, we simplified the model. To this end, instead of a fully latent structural regression model we conducted a partially latent structural regression model, in which the predictor variables were modelled as single indicators while controlling for measurement error [@Kline.2016, p. 214]. To get high-quality single indicators of the predictors, we computed the average of the model predicted values / latent factor scores, which we extracted from the CFAs. If the CFAs showed a unidimensional solution, we used the model predicted values for this latent factor; if the CFAs produced a multidimensional solution, we used the model predicted values for the general latent factor (again, by using this procedure we were able to attain a general measure of the predictors, instead of needing to further complicate analyses by including subdimensions of the predictors.) As the main variable of interest, the outcome need for privacy was measured as a three-dimensional latent factor consisting of overall 12 items (see below).

Finally, combining several items into a latent factors helps to reduce and condense information. At the same time, however, this approach also introduces several researcher degrees of freedom. We hence emphasize that we deleted items only in order to achieve satisfactory factorial validity and not in order to cherry-pick significant material. In light of our not having preregistered the analyses and in order to provide the complete picture, in the online supplementary material (OSM) we hence also report how the independent variables predicted each item individually.

`r cite_r("../bibliography/r-references.bib", withhold = FALSE, pkgs = c("GGally", "ggplot2", "knitr", "lavaan", "mice", "MVN", "papaja", "PerFit", "psych", "pwr", "semTools", "tidyverse"))` were used for the analyses.

@@

As the main variable of interest, we collected several items to measure need for privacy. First, we collected four items capturing the general need for privacy using the subscale _information_ developed by @Trepte.2017d. Second, we measured need for privacy on a societal level using nine self-designed items. The first subdimension was government surveillance, which represents the extent to which people want the government to abstain from collecting information about them. The second dimension was anonymity, which captures the extent to which people feel the need to avoid identification. Third, we measured need for privacy on an interpersonal level using nine self-designed items. The first subdimension measures need for privacy from other people in online contexts. The second subdimension measures need for privacy from other people in offline contexts.

EFAs and CFAs revealed a well-fitting model with 12 items (one double-loading), which consisted of the three dimensions (a) need for privacy from the government, (b) need for privacy from other people, and (c) need for anonymity. For a list of all items, see Table \@ref(tab:privacyitems).

```{r privacyitems}
d_t_items <- rbind(
  t_pri_nee %>% filter(name %in% pri_nee_gov_items),
  t_pri_nee %>% filter(name %in% pri_nee_int_items),
  t_pri_nee %>% filter(name %in% pri_nee_ano_items)) %>% 
  select("Item content" = content) %>% 
  mutate_all(linebreak)
kable(d_t_items, "latex", booktabs = TRUE, linesep = "",
      caption = "Overview of Items Measuring Need for Privacy") %>% 
  # column_spec(1, width = "16cm") %>% 
  kable_styling(latex_options="scale_down") %>% 
  kableExtra::group_rows("Privacy need government", 1, 5) %>%
  kableExtra::group_rows("Privacy need interpersonal", 6, 9) %>%
  kableExtra::group_rows("Privacy need anonymity", 10, 13)
```

@@

Sociability was measured with the Extraversion subscale Gregariousness [@Costa.1992b], which consists of eight items. An example item is “I shy away from crowds of people.” Analyses revealed that a bifactor model based on a subset of six items showed good fit to the data.

@@

Anxiety was measured with the Neuroticism subscale Worry [@Costa.1992b], which consists of eight items. An example item is “I am easily frightened.” Analyses revealed that a bifactor model using all eight items showed good fit to the data.
